---
title: "Module 3: Semi-supervised learning (PheCAP)"
author: "Siyue Yang, Jianhui Gao, and Jessica Gronsbell"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Load the packages. 
packages <- c("tidyverse", "PheCAP", "glmnet", "glmpath", "pROC")

# Check if the packages are missing or not. 
# If missing, install automatically.
# If not missing, load the package.
package.check <- lapply(
	  packages,
	  FUN = function(x) {
	    if (!require(x, character.only = TRUE)) {
	      install.packages(x, dependencies = TRUE)
	      library(x, character.only = TRUE)
	    }
	  }
	)
```

```{r}
# Load helper functions.
source("../Rscripts/helper_function.R")
```

```{r}
load('environment_pass.RData')
```

# Feature selection

## How to select features? 

Can leverage some clinical-meaningful features that are related to Y.

e.g. Feature "surrogate" = the total number of the disease-related billing codes + disease-specific NLP mentions.

```{r}
cbind(label = y, x) %>%
  filter(!is.na(label)) %>%
  mutate(disease = factor(label)) %>%
  ggplot(aes(x = surrogate)) +
  geom_density(aes(color = disease))
```

The more the disease-related codes, the more **likely** the patient has the disease.

```{r, warning=FALSE}
nonmissing_index <- which(!is.na(y))
surrogate <- x$surrogate

get_auc(y[nonmissing_index], surrogate[nonmissing_index])
```

We call these highly predictive features of the true disease status "surrogates".

## Opportunities of using surrogate features

1. Feature selection to reduce $p$

2. Algorithm development with limited $Y$

3. Algorithm validation with limited $Y$

Opportunity 2 and 3 will be covered in the next module! 

# Feature selection method

Motivation (Extreme assumption): 

- Patients with **high** main ICD or NLP mentions generally have the phenotype.
- Patients with **extremely** low counts are unlikely to have the phenotype.


```{r}
cbind(label = y, x) %>%
  filter(!is.na(label)) %>%
  mutate(disease = factor(label)) %>%
  ggplot(aes(x = surrogate)) +
  geom_density(aes(color = disease)) +
  theme_bw() + 
  annotate("rect", fill = "grey", alpha = 0.7, xmin = -1, xmax = 2, 
           ymin = -Inf, ymax = Inf)
```

- Left white rect: patients not having the disease.
- Right white rect: patients having the disease. 

## Prepare data for feature selection

### Prepare surrogates

Surrogates are available for all the patients!

```{r}
# Prepare 3 surrogates.
surrogate <- x$surrogate

# Prepare features to be selected. 
features <- data.matrix(x %>% select(starts_with("COD") | starts_with("NLP")))
```

Run surrogate-assisted feature extraction (SAFE) and show result.

```{r}
# Truncated at 2 and -1. 
SAFE <- extreme_method(surrogate, features, u_bound = 2, l_bound = -1)
SAFE_feature <- colnames(features)[SAFE$beta_select]
SAFE_feature
```

We select features that occur 50% among the three different surrogate-selected feature sets.
This is the idea of *majority voting*. 

Train phenotyping model and show the AUC on the testing set. 

-   Split data into training and testing set
-   Training 60% (n = 106), Testing 40% (n = 75)

```{r}
selected_features <- c("surrogate", "healthcare_utilization", SAFE_feature)

beta.phecap <- adaptive_lasso_fit(x = train_x[, selected_features], 
                                  y = train_y, 
                                  tuning = "cv", family = "binomial")

y_hat.phecap <- linear_model_predict(beta = beta.phecap, 
                                     x = test_x[, selected_features],
                                     probability = TRUE)
```

```{r}
roc.lasso <- roc(test_y, y_hat.lasso)
roc.alasso <- roc(test_y, y_hat.alasso)
roc.phecap <- roc(test_y, y_hat.phecap)

plot(roc.lasso,
  print.auc = TRUE, main = "n_training = 106 (40%)"
)
plot(roc.alasso,
  print.auc = TRUE, col = 'red', add = TRUE, print.auc.y = 0.4
)
plot(roc.phecap,
  print.auc = TRUE, col = 'blue', add = TRUE, print.auc.y = 0.3
)
legend(0, 0.2, legend = c("LASSO", "ALASSO", "PheCAP"), col = c("black","red"), 
       lty = 1, cex = 0.8)
```

```{r}
roc_full.lasso <- get_roc(y_true = test_y, y_score = y_hat.lasso)
head(roc_full.lasso,10)
```

```{r}
roc_full.alasso <- get_roc(y_true = test_y, y_score = y_hat.alasso)
head(roc_full.lasso,10)
```

```{r}
roc_full.phecap <- get_roc(y_true = test_y, y_score = y_hat.phecap)
head(roc_full.phecap,10)
```


Save the data and feature selected for module 4 and model fitting. 

```{r, eval = F}
save(list = ls(), file = "../module4/environment_pass.RData")
```
