---
title: "Module 3: Semi-supervised learning (PheCAP)"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, include=FALSE}
# Load the packages. 
packages <- c("tidyverse", "PheCAP", "glmnet", "glmpath", 
              "pROC", "parallel", "ggpubr")

# Check if the packages are missing or not. 
# If missing, install automatically.
# If not missing, load the package.
package.check <- lapply(
	  packages,
	  FUN = function(x) {
	    if (!require(x, character.only = TRUE)) {
	      install.packages(x, dependencies = TRUE)
	      library(x, character.only = TRUE)
	    }
	  }
	)
```

```{r, include=FALSE}
# Load helper functions.
source("../Rscripts/helper_function.R")

# Load data. 
load("environment.RData")
```

# Surrogates for CAD

```{r, include = FALSE}
icd <- ehr_data %>%
  filter(!is.na(label)) %>%
  mutate(disease = factor(label)) %>%
  ggplot(aes(x = log(main_ICD + 1))) +
  geom_density(aes(color = disease))

nlp <- ehr_data %>%
  filter(!is.na(label)) %>%
  mutate(disease = factor(label)) %>%
  ggplot(aes(x = log(main_NLP + 1))) +
  geom_density(aes(color = disease))
```

```{r, fig.height=3, echo=FALSE}
figure <- ggarrange(icd, nlp,
                    labels = c("ICD", "NLP"),
                    ncol = 2, nrow = 1)
figure
```

The more the disease-related codes and NLP mentions, the more **likely** the patient has the disease.

# ROC Surrogates

\tiny
```{r, warning=FALSE, include = FALSE}
nonmissing_index <- which(!is.na(ehr_data$label))
y <- ehr_data$label[nonmissing_index]
sicd <- ehr_data$main_ICD[nonmissing_index]
snlp <- ehr_data$main_NLP[nonmissing_index] 

roc_icd <- roc(y, log(sicd + 1))
roc_nlp <- roc(y, log(snlp + 1))

# Prepare features to be selected. 
x <- data.matrix(ehr_data %>% select(starts_with("COD") | starts_with("NLP")))
```

```{r, echo=FALSE}
ggroc(list(roc_icd, roc_nlp), legacy.axes = TRUE) +
  scale_color_discrete(labels = c(
    paste("ICD, AUC =", round(auc(roc_icd), 3)),
    paste("NLP, AUC =", round(auc(roc_nlp), 3))
  )) + theme(legend.position = "bottom") +
  ggtitle("The operating receiver characteristic (ROC) curve") + 
  guides(color = guide_legend(ncol = 2, title = element_blank())) + 
  labs(x = "False positive rate (FPR)", y = "True positive rate (TPR)")
```

# Step 1: SAFE 

\tiny
```{r, cache=TRUE}
surrogates <- list(
  PhecapSurrogate(
    variable_names = "main_ICD",
    lower_cutoff = 1, upper_cutoff = 10),
  PhecapSurrogate(
    variable_names = "main_NLP",
    lower_cutoff = 1, upper_cutoff = 10))

feature_selected <- phecap_run_feature_extraction(data, surrogates)
feature_selected
```

# Step 2: Orthogonalization + supervised learning

\tiny
```{r}
phecap_lasso <- phecap_train_phenotyping_model(data, surrogates, feature_selected,
                                        method = "lasso_cv")
```

\tiny
```{r}
phecap_lasso
```

```{r, echo=FALSE}
# Get model predictions + ROC curve.
phenotype <- phecap_predict_phenotype(data, phecap_lasso)
y_hat_phecap <- phenotype[test_data$patient_id, "prediction"]
roc_phecap <- roc(test_y, y_hat_phecap)
```

# Supervised learning (LASSO, ALASSO) vs. PheCAP 

```{r, echo = FALSE}
ggroc(list(roc_lasso, roc_alasso, roc_phecap), legacy.axes = TRUE) +
  scale_color_discrete(labels = c(
    paste("LASSO, AUC =", round(auc(roc_lasso), 3)),
    paste("ALASSO, AUC =", round(auc(roc_alasso), 3)),
    paste("PheCAP, AUC =", round(auc(roc_phecap), 3))
  )) + theme(legend.position = "bottom") +
  ggtitle("The operating receiver characteristic (ROC) curve") + 
  guides(color = guide_legend(ncol = 2, title = element_blank())) + 
  labs(x = "False positive rate (FPR)", y = "True positive rate (TPR)")
```

# Supervised learning vs. PheCAP for different training size

- Randomly sample training size $=50,70,90$
- Use the remaining data as the test set
- Repeat 600 times

\tiny
```{r, cache=TRUE}
auc_phecap <- validate_phecap(dat = labeled_data, 
                              orig_data = ehr_data,
                              surrogates = surrogates,
                              feature_selected = feature_selected,
                              nsim = 600,
                              n.train = c(50, 70, 90))
```

# Supervised learning vs. PheCAP for different training size

```{r, echo = FALSE}
tidyr::gather(cbind(auc_supervised, auc_phecap)) %>%
  mutate(
    n = gsub(",.*$", "", key),
    method = sub(".*,\\s*", "", key)
  ) %>%
  ggplot(aes(y = value, color = method)) +
  geom_boxplot(outlier.shape = NA) +
  facet_wrap(. ~ n) +
  ggtitle("Area under the ROC curve (AUC) from 600 simulations") +
  theme(legend.position = "bottom", axis.text.x = element_blank(), axis.ticks = element_blank())
```

```{r, echo = FALSE, eval = FALSE}
save(list = ls(), file = "../module4/environment.RData")
rmarkdown::render("module3_slides.Rmd")
```

