---
title: "Module 3: Semi-supervised learning (PheCAP)"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, include=FALSE}
# Load the packages. 
packages <- c("tidyverse", "PheCAP", "glmnet", "glmpath", 
              "pROC", "parallel", "ggpubr")

# Check if the packages are missing or not. 
# If missing, install automatically.
# If not missing, load the package.
package.check <- lapply(
	  packages,
	  FUN = function(x) {
	    if (!require(x, character.only = TRUE)) {
	      install.packages(x, dependencies = TRUE)
	      library(x, character.only = TRUE)
	    }
	  }
	)
```

```{r, include=FALSE}
# Load helper functions.
source("../Rscripts/helper_function.R")

# Load data. 
load("../module3/environment.RData")
```

# Surrogates for CAD

```{r, include = FALSE}
icd <- ehr_data %>%
  filter(!is.na(label)) %>%
  mutate(disease = factor(label)) %>%
  ggplot(aes(x = log(main_ICD + 1))) +
  geom_density(aes(color = disease)) +
  theme(text = element_text(size = 15)) +
  ggtitle("Density plot of the main ICD codes")

nlp <- ehr_data %>%
  filter(!is.na(label)) %>%
  mutate(disease = factor(label)) %>%
  ggplot(aes(x = log(main_NLP + 1))) +
  geom_density(aes(color = disease)) +
  theme(text = element_text(size = 15)) +
  ggtitle("Density plot of the main NLP codes")
```

```{r, fig.height=4, echo=FALSE}
figure <- ggarrange(icd, nlp, ncol = 2, nrow = 1, common.legend = TRUE)
figure
```

The more the disease-related codes and NLP mentions, the more **likely** the patient has the disease.

# ROC Surrogates

\tiny
```{r, warning=FALSE, include = FALSE}
nonmissing_index <- which(!is.na(ehr_data$label))
y <- ehr_data$label[nonmissing_index]
sicd <- ehr_data$main_ICD[nonmissing_index]
snlp <- ehr_data$main_NLP[nonmissing_index] 

roc_icd <- roc(y, log(sicd + 1))
roc_nlp <- roc(y, log(snlp + 1))

# Prepare features to be selected. 
x <- data.matrix(ehr_data %>% select(starts_with("COD") | starts_with("NLP")))
```

```{r, echo=FALSE}
plot_roc(list(roc_icd, roc_nlp), legend = c("ICD", "NLP"))
```

# Step 1: SAFE 

\tiny
```{r, cache=TRUE}
surrogates <- list(
  PhecapSurrogate(
    variable_names = "main_ICD",
    lower_cutoff = 1, upper_cutoff = 10),
  PhecapSurrogate(
    variable_names = "main_NLP",
    lower_cutoff = 1, upper_cutoff = 10))

feature_selected <- phecap_run_feature_extraction(data, surrogates)
feature_selected
```

# Step 2: Orthogonalization + supervised learning

\tiny
```{r}
phecap_lasso <- phecap_train_phenotyping_model(data, surrogates, feature_selected,
                                        method = "lasso_cv")
```

\tiny
```{r}
phecap_lasso
```

```{r, echo=FALSE}
# Get model predictions + ROC curve.
phenotype <- phecap_predict_phenotype(data, phecap_lasso)
y_hat_phecap <- phenotype[test_data$patient_id, "prediction"]
roc_phecap <- roc(test_y, y_hat_phecap)
```


# Supervised learning (LASSO, ALASSO) vs. PheCAP 

```{r, echo = FALSE}
plot_roc(list(roc_lasso, roc_alasso, roc_phecap), 
         legend = c("LASSO", "ALASSO", "PheCAP"),
         method_index = c(1,2,5))
```

# Supervised learning (LASSO, ALASSO) vs. PheCAP at FPR = 0.10

\tiny
```{r}
get_roc_parameter(0.1, roc_full_lasso)
get_roc_parameter(0.1, roc_full_alasso)

roc_full_phecap <- get_roc(y_true = test_y, y_score = y_hat_phecap) %>% data.frame()
get_roc_parameter(0.1, roc_full_phecap)
```

# Supervised learning vs. PheCAP for different training size

- Randomly sample training size $=50,70,90$
- Use the remaining data as the test set
- Repeat 600 times

\tiny
```{r, cache=TRUE}
auc_phecap <- validate_phecap(dat = labeled_data, 
                              surrogates = surrogates,
                              feature_selected = feature_selected,
                              nsim = 600,
                              n.train = c(50, 70, 90))
```

# Supervised learning vs. PheCAP for different training size

```{r, echo = FALSE}
plot_sims(cbind(auc_supervised, auc_phecap), 
          legend = c("LASSO", "ALASSO", "PheCAP"),
          method_index = c(1,2,5))
```

```{r, echo = FALSE, eval = FALSE}
save(list = ls(), file = "../module4/environment.RData")
rmarkdown::render("module3_slides.Rmd")
```

