---
title: "Electronic Health Records Phenotyping"
description: |
  SSC Annual Meeting 2022 Biostatistics Workshop
author: 
- name: Jessica Gronsbell, Ph.D.
  url: https://sites.google.com/view/jgronsbell/home?authuser=0
  affiliation: University of Toronto
  affiliation_url: https://www.statistics.utoronto.ca/
- name: Siyue Yang
- name: Jianhui Gao
date: June 5th, 2022
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages
```

# Welcome

This website contains materials for Biostatistics Workshop in [2022 SSC Annual Meeting Online](https://ssc.ca/en/meetings/annual/2022-annual-meeting). 

In this workshop, we will introduce statistical learning methods designed to expedite the phenotyping process in order to improve the scalability of EHR research. 

- **Module 1** Exploratory Data Analysis (EDA) and Feature Selection ([Exerciese 1](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module1/ex1.Rmd))

- **Module 2** Model Fitting and Evaluation  ([Exerciese 2](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module2/ex2.Rmd))


### PheCAP

We will use the real EHR data examples from [PheCAP](https://celehs.github.io/PheCAP/), an R package that implements high-throughput phenotyping with electronic health record (EHR) using a common automated pipeline.. 

### Required packages

Here is a list of packages that will be used during the workshop. 

```{r, eval=FALSE, echo=TRUE}
# From CRAN
install.packages("PheCAP")
install.packages("PheNorm")
install.packages("MAP")

install.packages("tidyverse")
install.packages("corrplot")

install.packages("glmnet")
install.packages("randomForestSRC")
install.packages("pROC")
install.packages("mclust")
```


# Resources

- Publicly available EHR dataset

  - Physionet 
  - n2n2 competitions
  
- NLP software
  - cTAKES
  - clinspacy
  - medspacy
  

# Acknowledgments {.appendix}

This website was made with the [distill package](https://rstudio.github.io/distill/), drawn mostly from Silvia Canelón’s [Sharing Your Work with xaringan](https://spcanelon.github.io/xaringan-basics-and-beyond/). 



