---
title: "Electronic Health Records Phenotyping"
description: |
  SSC Annual Meeting 2022, Biostatistics Workshop
author: 
- name: Jesse Gronsbell, Ph.D.
  url: https://sites.google.com/view/jgronsbell/home?authuser=0
  affiliation: University of Toronto
  affiliation_url: https://www.statistics.utoronto.ca/
- name: Jianhui Gao
- name: Siyue Yang
date: June 5th, 2022
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages
```

# Welcome

This website contains materials for Biostatistics Workshop at the [2022 SSC Annual Meeting Online](https://ssc.ca/en/meetings/annual/2022-annual-meeting). 

In this workshop, we introduce methods for phenotyping with electronic health record (EHR) data.  Slides and code can be found for an example phenotyping problem below.

- Electronic health records phenotyping [[Slides]()]

- **Module 1** Introduction [[Slides](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module1/module1_slides.pdf)] [[Rmd](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module1/module1_slides.Rmd)]

- **Module 2** Supervised learning [[Slides](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module2/module2_slides.pdf)] [[Rmd](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module2/module2_slides.Rmd)]

- **Module 3** Semi-supervised learning (PheCAP)  [[Slides](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module3/module3_slides.pdf)] [[Rmd](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module3/module3_slides.Rmd)]

- **Module 4** Alternative approaches  [[Slides](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module4/module4_slides.pdf)] [[Rmd](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module4/module4_slides.Rmd)]



### PheCAP

We use real EHR data from [PheCAP](https://celehs.github.io/PheCAP/), an R package that implements high-throughput phenotyping using a common automated pipeline. 

### Required packages

Please copy and paste the following code into R Console to check and load the packages. We are using R version $\geq$ 3.6.0.

```{r, eval=FALSE, echo=TRUE}
# Specify the packages from CRAN.
packages <- c(
	  "PheCAP",     
	  "PheNorm",  
	  "MAP", 
	  "glmnet",     
	  "tidyverse",  
	  "ggplot2",  
	  "data.table", 
	  "mltools",  
	  "pROC",
	  "parallel",
	  "randomForestSRC",
	  "SVMMaj"
)

# Load the packages.
# Missing package(s) will be first installed and loaded.  
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

### Implementation

To start with, please `git clone` the repository to local. 

```{r, eval=FALSE, echo=TRUE}
git clone git@github.com:jlgrons/EHR-Phenotyping-Workshop.git
```


# Resources

- Publicly available EHR dataset

  - [PhysioNet](https://physionet.org/)
  - [National NLP Clinical Challenges (n2c2)](https://n2c2.dbmi.hms.harvard.edu/) NLP competitions
  
- NLP software
  - [cTAKES](https://ctakes.apache.org/)
  - [clinspacy](https://github.com/ML4LHS/clinspacy)
  - [medspacy](https://github.com/medspacy/medspacy)
  

# Acknowledgments {.appendix}

This website was made with the [distill package](https://rstudio.github.io/distill/), drawn mostly from Silvia Canelón’s [Sharing Your Work with xaringan](https://spcanelon.github.io/xaringan-basics-and-beyond/). 
