---
title: "Electronic Health Records Phenotyping"
description: |
  SSC Annual Meeting 2022 Biostatistics Workshop
author: 
- name: Jessica Gronsbell, Ph.D.
  url: https://sites.google.com/view/jgronsbell/home?authuser=0
  affiliation: University of Toronto
  affiliation_url: https://www.statistics.utoronto.ca/
- name: Siyue Yang
- name: Jianhui Gao
date: June 5th, 2022
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages
```

# Welcome

This website contains materials for Biostatistics Workshop in [2022 SSC Annual Meeting Online](https://ssc.ca/en/meetings/annual/2022-annual-meeting). 

In this workshop, we will introduce statistical learning methods designed to expedite the phenotyping process in order to improve the scalability of EHR research. 

- **Module 1** Exploratory Data Analysis (EDA) and Feature Selection ([Exerciese 1](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module1/ex1.Rmd))

- **Module 2** Model Fitting and Evaluation  ([Exerciese 2](https://github.com/jlgrons/EHR-Phenotyping-Workshop/blob/main/module2/ex2.Rmd))


### PheCAP

We will use the real EHR data examples from [PheCAP](https://celehs.github.io/PheCAP/), an R package that implements high-throughput phenotyping with electronic health record (EHR) using a common automated pipeline. 

### Required packages

Here is a list of packages that will be used during the workshop. 

Please copy and paste the following code into R Console to check and load the packages. 

We are using R version ($\geq$ 3.6.0).

```{r, eval=FALSE, echo=TRUE}
# Specify the packages from CRAN.
packages <- c(
	  "PheCAP",     
	  "PheNorm",   
	  "MAP",        
	  "glmnet",     
	  "glmpath",  
	  "mclust",     
	  "tidyverse",  
	  "ggplot2",  
	  "corrplot",
	  "data.table", 
	  "mltools",  
	  "pROC",
	  "randomForestSRC"
)

# Load the packages.
# Missing package(s) will be first installed and loaded.  
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```


# Resources

- Publicly available EHR dataset

  - Physionet 
  - n2n2 competitions
  
- NLP software
  - cTAKES
  - clinspacy
  - medspacy
  

# Acknowledgments {.appendix}

This website was made with the [distill package](https://rstudio.github.io/distill/), drawn mostly from Silvia Canelón’s [Sharing Your Work with xaringan](https://spcanelon.github.io/xaringan-basics-and-beyond/). 
