---
title: "Module 4: Alternative approaches"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
# If a package is installed, it will be loaded. If any
## are not, the missing package(s) will be installed
## from CRAN and then loaded.

## First specify the packages of interest
packages <- c(
  "dplyr", "PheCAP", "glmnet", "randomForestSRC", "PheNorm",
  "MAP", "pROC", "mltools", "data.table", "ggplot2", "parallel"
)

## Now load or install&load all
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

# load environment from example 1
load("../module4/environment.RData")
source("../Rscripts/helper_function.R")
```


# 2-step Semi-supervised Approach
i) Regress the surrogate on the features with penalized least square to get the direction of beta.

```{r, cache = TRUE}
x <- all_x %>% select(starts_with("health") | starts_with("COD") | starts_with("NLP")) # COD + NLP
S <- ehr_data$main_ICDNLP

# Step 1
beta.step1 <- adaptive_lasso_fit(
  y = S, # surrogate
  x = x, # all X
  family = "gaussian",
  tuning = "cv"
)

# Features selected
names(beta.step1[abs(beta.step1) > 0])[-1]
```

# 2-step Semi-supervised Approach
i) Regress the surrogate on the features with penalized least square to get the direction of beta.

(ii) Regress the outcome on the linear predictor to get the intercept and multiplier for the beta.
```{r}
# linear predictor without intercept
bhatx <- linear_model_predict(beta = beta.step1, x = as.matrix(x))

# Step 2
step2 <- glm(train_y ~ bhatx[train_data$patient_id] + S[train_data$patient_id] +
  health_count[train_data$patient_id])
beta_step2 <- coef(step2)
beta_step2
```

```{r}
# recover beta
beta <- beta_step2[2] * beta.step1
# mu
mu <- beta_step2[1] +
  as.numeric(as.matrix(x[test_data$patient_id, ]) %*% beta[-1]) +
  as.numeric(beta_step2[3] %*% S[test_data$patient_id]) +
  as.numeric(beta_step2[4] %*% health_count[test_data$patient_id])
# expit
y_hat.ss <- plogis(mu)
```


# ROC
```{r}
plot(roc(test_y, y_hat.lasso),
  print.auc = TRUE, main = "n_training = 106 (60%)"
)
plot(roc(test_y, y_hat.alasso),
  print.auc = TRUE, col = "red", add = TRUE, print.auc.y = 0.4
)
plot(roc(test_y, y_hat.ss),
  print.auc = TRUE, col = "green", add = TRUE, print.auc.y = 0.2
)
plot(roc(test_y, y_hat.phecap),
  print.auc = TRUE, col = "blue", add = TRUE, print.auc.y = 0.3
)
legend(0, 0.3,
  legend = c("LASSO", "ALASSO", "PheCAP", "Two-Step"),
  col = c("black", "red", "blue", "green"),
  lty = 1, cex = 0.8
)
```

# Model Evaluation
```{r, cache=TRUE}
start <- Sys.time()
auc_twostep <- validate_ss(
  dat = labeled_data, nsim = 600,
  n.train = c(50, 70, 90),
  beta = beta.step1,
  S = S,
  x = x
)
end <- Sys.time()
end - start
```




```{r}
par(mfrow = c(1,3))
# Compare with Previous method
boxplot(cbind(auc_supervised, auc_phecap, auc_twostep)
%>% select(starts_with("n=50")),
ylim = c(0.5, 1), names = c("LASSO", "ALASSO","PheCAP", "Two-Step")
, main = "n=50"
)
boxplot(cbind(auc_supervised, auc_phecap, auc_twostep)
%>% select(starts_with("n=70")),
ylim = c(0.5, 1), names = c("LASSO", "ALASSO","PheCAP", "Two-Step"), 
main = "n=70"
)
boxplot(cbind(auc_supervised, auc_phecap, auc_twostep)
%>% select(starts_with("n=90")),
ylim = c(0.5, 1), names = c("LASSO", "ALASSO","PheCAP","Two-Step"), 
main = "n=90"
)
```

## MAP
```{r}
# Use untransformed data; MAP requires sparse matrix
# Create sparse matrix for surroagtes
data_fit <- sparsify(PheCAP::ehr_data %>%
  select(main_ICD, main_NLP) %>%
  rename(ICD = main_ICD) %>% data.table())

# Create sparse matrix for HU
note <- Matrix(PheCAP::ehr_data$healthcare_utilization, ncol = 1, sparse = TRUE)
model_map <- MAP(mat = data_fit, note = note, full.output = TRUE)

y_hat.map <- model_map$scores[data$validation_set]
```

```{r}
plot(roc(test_y, y_hat.lasso),
  print.auc = TRUE, main = "n_training = 106 (60%)"
)
plot(roc(test_y, y_hat.alasso),
  print.auc = TRUE, col = "red", add = TRUE, print.auc.y = 0.4
)
plot(roc(test_y, y_hat.phecap),
  print.auc = TRUE, col = "blue", add = TRUE, print.auc.y = 0.3
)
plot(roc(test_y, y_hat.ss),
  print.auc = TRUE, col = "green", add = TRUE, print.auc.y = 0.2
)
plot(roc(test_y, y_hat.map),
  print.auc = TRUE, col = "orange", add = TRUE, print.auc.y = 0
)

legend(0, 0.4,
  legend = c("LASSO", "ALASSO", "PheCAP","Two-Step",  "PheNorm", "MAP"),
  col = c("black", "red", "blue", "green", "orange"),
  lty = 1, cex = 0.8
)
```