---
title: "Module 4: Alternative approaches"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
# If a package is installed, it will be loaded. If any
## are not, the missing package(s) will be installed
## from CRAN and then loaded.

## First specify the packages of interest.
packages <- c(
  "dplyr", "PheCAP", "glmnet", "randomForestSRC", "PheNorm",
  "MAP", "pROC", "mltools", "data.table", "ggplot2", "parallel"
)

## Now load or install&load all.
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

# Load environment from example 1.
load("../module4/environment.RData")
source("../Rscripts/helper_function.R")
```

# 2-step Semi-supervised Approach

1. Regress the surrogate on the features with penalized least square to get 
the direction of beta.

```{r}
# COD + NLP + HU.
x <- log(ehr_data %>% select(starts_with("health") | 
  starts_with("COD") | starts_with("NLP")) + 1) 
S <- log(ehr_data$main_ICD + ehr_data$main_NLP + 1)

# Step 1.
beta_step1 <- adaptive_lasso_fit(
  y = S, # surrogate
  x = x, # all X
  family = "gaussian",
  tuning = "cv"
)
```

# 2-step Semi-supervised Approach

1. Regress the surrogate on the features with penalized least square to get the direction of beta.

2. Regress the outcome on the linear predictor to get the intercept and multiplier for the beta.

\tiny 
```{r}
# Linear predictor without intercept.
bhatx <- linear_model_predict(beta = beta_step1, x = as.matrix(x))

# Step 2.
step2 <- glm(
  train_y ~ bhatx[train_data$patient_id] + S[train_data$patient_id],
  family = "binomial"
)
beta_step2 <- coef(step2)
beta_step2

# Recover beta.
beta <- beta_step2[2] * beta_step1
```

# Compare selected features

\tiny

```{r}
# LASSO.
names(beta_lasso[!beta_lasso == 0])[-1]
```

```{r}
# ALASSO.
names(beta_alasso[!beta_alasso == 0])[-1]
```

# Compare selected features

\tiny 

```{r}
# PheCAP.
feature_selected
```

```{r}
# Two Step.
names(beta[!beta == 0])[-1]
```

# ROC

```{r}
mu <- beta_step2[1] +
  as.numeric(as.matrix(x[test_data$patient_id, ])
  %*% beta[-1]) +
  as.numeric(beta_step2[3] %*% S[test_data$patient_id])

# Expit.
y_hat_twostep <- plogis(mu)

roc_twostep <- roc(test_y, y_hat_twostep)
```

# ROC

```{r, echo=FALSE}
plot_roc(
  list(roc_lasso, roc_alasso, roc_phecap, roc_twostep),
  legend = c("LASSO", "ALASSO", "PheCAP", "Two-step"),
  method_index = c(1,2,5,6)
)
```

# Model Evaluation

```{r, cache=TRUE, echo=FALSE}
auc_twostep <- validate_ss(
  dat = labeled_data, nsim = 600,
  ntrain = c(50, 70, 90),
  beta = beta_step1,
  S = S,
  x = x
)
```


```{r, echo=FALSE}
plot_sims(
  cbind(auc_supervised, auc_phecap, auc_twostep),
  legend = c("LASSO", "ALASSO", "PheCAP", "Two-step"),
  method_index = c(1,2,5,6)
)
```

# MAP

\tiny 
```{r, message=FALSE}
# Use un-transformed data; MAP requires sparse matrix.
# Create sparse matrix for surrogates.
data_fit <- sparsify(
  PheCAP::ehr_data %>%
  select(main_ICD, main_NLP) %>%
  rename(ICD = main_ICD) %>% data.table()
)

# Create sparse matrix for HU.
note <- Matrix(
  PheCAP::ehr_data$healthcare_utilization,
  ncol = 1, sparse = TRUE
)
model_map <- MAP(mat = data_fit, note = note, full.output = TRUE)

y_hat_map <- model_map$scores[data$validation_set]
roc_map <- roc(test_y, y_hat_map)
```

# ROC

```{r, echo=FALSE}
plot_roc(
  list(roc_lasso, roc_alasso, roc_phecap, roc_twostep, roc_map),
  legend = c("LASSO", "ALASSO", "PheCAP", "Two-step", "MAP"),
  method_index = c(1,2,5,6,7)
)
```
