---
title: "Module 1: Introduction"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# CAD data overview

This data mart is a random sample of patients in the Mass General Brigham (formerly Partner's Healthcare) EHR database who had at least one note of 500 characters and met an initial filter for **coronary artery disease** (CAD) defined as: 
    
- $\geq$ 1 ICD9 code related to CAD (410.x, 411.x, 412.x, 413.x, 414.x). 
- $\ge$ 1 mention for any CAD related concepts (eg. CAD, CAD procedures, CAD biomarkers, positive stress test).

```{r, echo=F}
# Packages required for this module.
packages <- c("tidyverse", "PheCAP", "corrplot", "ggplot2")

# Load packages, or install if missing.
package.check <- lapply(
	  packages,
	  FUN = function(x) {
	    if (!require(x, character.only = TRUE)) {
	      install.packages(x, dependencies = TRUE)
	      library(x, character.only = TRUE)
	    }
	  }
	)
```

```{r, echo=F}
# Load helper functions.
source("../Rscripts/helper_function.R")
```

# Read in the CAD data

"label": whether the patient has CAD, **extracted from chart review by a clinician**.

\tiny
```{r}
data(ehr_data)
data <- PhecapData(
  ehr_data, "healthcare_utilization", "label", 75, 
  patient_id = "patient_id", seed = 123
)
data
```

# Read in the CAD data

"main_ICD", "main_NLP": total number of billing codes or NLP mentions of the disease.

\tiny
```{r}
head(ehr_data[, c(1:5, 25:30)])
```

# Read in the CAD data

"healthcare_utilization": total number of notes the patient has.

\tiny
```{r}
head(ehr_data[, c(1:5, 25:30)])
```

# Read in the CAD data

“CODx” (n = 10): the counts of a specific code.
“NLPx” (n = 574): the counts of a NLP term.

\tiny
```{r}
head(ehr_data[, c(1:5, 25:30)])
```

# Basic descriptives

```{r}
# Check for missing data.
colnames(ehr_data)[which(colMeans(is.na(ehr_data)) > 0)]
```

\pause

```{r}
# Prevalence of the label.
mean(ehr_data$label, na.rm = TRUE)
```

# Feature distributions

```{r echo=FALSE}
set.seed(99)
feature_index <- sample(c(3:ncol(ehr_data)), 10, replace = FALSE)
ehr_data[, feature_index] %>%
  pivot_longer(everything(), names_to = "feature", values_to = "count") %>%
  ggplot() +
  geom_density(aes(x = count, color = feature))+
  theme(text = element_text(size = 20)) +
  ggtitle("Density plots of randomly selected 10 features")
```

# Prepare the data for model fitting

- We log transform all the features as they are highly skewed

- We orthogonalize all features against health care utilization before fitting as patients with higher healthcare utilization have higher feature counts (in Module 3)

(Please find more details in the `.Rmd` file.)

```{r, echo=FALSE}
# Log transformed the features. 
features <- ehr_data[, c(3:ncol(ehr_data))]
features <- log(features + 1)

# Save the transformed features. 
# Please note that in the beginning of module 2, the transformation is also performed.
ehr_data_transformed <- cbind(
  patient_id = ehr_data$patient_id,
  label = ehr_data$label, 
  healthcare_utilization =features$healthcare_utilization, 
  features
) 
```

